---
title: "Assignment 1"
author: "Josh Bannon"
format: 
  html: default
  pdf: default
editor: visual
---

```{r echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(readr)
library(ggplot2)
library(knitr)
library(rmarkdown)
library(kableExtra)
```
## **Introduction**

This report investigates whether higher marketing spend per customer leads to increased purchase amounts in an online retail setting. The dataset represents a personalised marketing campaign where spending varied across customers through channels such as advertisements, emails, and coupons. Businesses often face uncertainty in determining how much to invest in marketing to maximise returns without overspending. The findings will help determine whether increased spending generates meaningful returns or leads to diminishing benefits.

## **Research Question**

This report examines whether higher marketing spend per customer leads to greater purchase amounts. Specifically, it investigates the relationship between the amount spent on personalised marketing efforts and the revenue generated from each customer, motivated by the need for businesses to optimise marketing investment decisions. The analysis aims to determine whether increased spending results in higher returns or diminishing effectiveness.

## **Dataset Introduction**

The dataset records customer-level information from a personalised marketing campaign, including marketing spend, purchase amount, and number of visits. The key variables of interest are marketing_spend, which represents the amount spent targeting each customer, and purchase_amount, which reflects the revenue generated from each customer. Additional variables such as visits, customer segment, and signup date provide further context on customer activity and characteristics but are not overly relevant to the research question. Before analysis, the dataset was cleaned by removing rows with missing values in purchase amount because we're measuring how higher marketing spend leads to higher purchase amount, we then standardised date formats as well as formatting numerical formats. Irrelivant variables such as segment and notes were excluded due to their insignificance and irrelivance

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Reading in the data
data <- read_csv("data/32592582_Joshua Bannon.csv")
```

```{r echo=FALSE, message=FALSE, warning=FALSE, eval=FALSE}
# Initial look at the data
glimpse(data)
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Removing rows with missing values in marketing_spend and purchase_amount
data <- data %>%
  drop_na(marketing_spend, purchase_amount)
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Standardising variable formats and removing irrelevant variables
data <- data %>%
  mutate(
    marketing_spend = as.numeric(marketing_spend),
    purchase_amount = as.numeric(purchase_amount),
    visits = as.numeric(visits),
    signup_date = as.Date(signup_date),
    segment = as.factor(segment)) %>% 
  select(-notes) %>%
  distinct()
```

### **Dataset Description subsection**

The dataset contains **`r nrow(data)`** observations and **`r ncol(data)`** variables, capturing customer-level information from a personalised marketing campaign. It includes a mix of numeric variables (e.g., marketing_spend, purchase_amount, visits), a categorical variable (segment), and a date variable (signup_date). Marketing spend ranges from **$`r min(data$marketing_spend)`** to **$`r max(data$marketing_spend)`**, while purchase amounts range from **$`r min(data$purchase_amount)`** to **$`r max(data$purchase_amount)`**. Customer visits range from **`r min(data$visits)`** to **`r max(data$visits)`**.

```{R}
data %>%
  group_by(segment) %>%
  summarise(
    average_spend = round(mean(marketing_spend), 2),
    average_purchase = round(mean(purchase_amount), 2),
    average_visits = round(mean(visits), 2),
    number_of_customers = n(),
    .groups = "drop"
  ) %>%
  as.data.frame()
```

```{r}
ggplot(data, aes(x = marketing_spend, y = purchase_amount)) +
  geom_point(colour = "steelblue") +
  geom_smooth(method = "lm", se = FALSE, colour = "red") +
  labs(
    title = "Relationship Between Marketing Spend & Purchase Amount",
    x = "Marketing Spend ($)",
    y = "Purchase Amount ($)"
  ) +
  theme_minimal() +
  theme(plot.title = element_text(face = "bold", hjust = 0.5)
  )
```

# **Results Section**

**Using visualisations of the data, discuss the answer to your research questions. You must:** **Create 2 figures maximum of your data that will help you answer your research question. Each figure must have a caption using the options inside the R code chunk. Create the figures using the ggplot2 package. The figures must have captions, and must be referred to in the text of the results section.**

```{r}
ggplot(data, aes(x = marketing_spend, y = purchase_amount)) +
  geom_point(alpha = 0.7) +
  geom_smooth(method = "lm", se = TRUE) +
  labs(
    x = "Marketing Spend",
    y = "Purchase Amount"
  )
```

```{r}
data <- data %>%
  mutate(
    spend_group = case_when(
      marketing_spend < quantile(marketing_spend, 0.33) ~ "Low",
      marketing_spend < quantile(marketing_spend, 0.67) ~ "Medium",
      TRUE ~ "High"
    )
  )
```

```{r}
ggplot(data, aes(x = spend_group, y = purchase_amount)) +
  geom_boxplot() +
  labs(
    x = "Marketing Spend Group",
    y = "Purchase Amount"
  )
```

**As well as the figures, you must also include a paragraph describing these results and how they link back to your research question. Remember, you can use inline code to extract certain numbers automatically.**
